---
title: "StudentTracker File Formatter"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    source_code: embed
---

```{r setup, include=FALSE}

# Code license: [MIT](https://www.r-project.org/Licenses/MIT

library(flexdashboard)
library(shiny)
library(dplyr)
library(stringr)
library(lubridate)

```

Column {data-width=300, .sidebar}
-----------------------------------------------------------------------
### Upload & Enter Data

```{r}
 
  

 fileInput('file1', 'Choose CSV File',
                accept=c('text/csv', 
                         'text/comma-separated-values,text/plain', 
                         '.csv'))
       radioButtons('sep', 'Separator',
                   c(Comma=',',
                      Semicolon=';',
                      Tab='\t'),
                    ',')
       radioButtons('quote', 'Quote',
                    c(None='',
                      'Double Quote'='"',
                      'Single Quote'="'"),
                    '"')

      textInput("institution_code", label = "Institution Code")
      textInput("branch_code", label = "Branch Code")
      textInput("institution_name", label = "Institution Name")
      textInput("inquiry_type", label = "Inquiry Type")
      downloadButton('downloadData', label = 'Download')
      

       downloadHandler(
      filename = function() { paste(input$institution_name, today(), '.txt', sep='') },
      content = function(file) {
        write.table(data(), file, col.names = FALSE, quote = FALSE, row.names = FALSE, sep =  "\t")
      }
    )


date_today <- today()

date_today <- paste0(year(date_today),
                     str_pad(month(date_today), 2, pad = "0"),
                     str_pad(day(date_today), 2, pad = "0"))
  
       data <- reactive({
      
      
      record_type <- input$record_type # creates record type
      institution_code <- input$institution_code # creates school code
      branch_code <- input$branch_code # creates branch code
      institution_name <- input$institution_name # creates school name
      date_create <- input$date_create # creates date create
      inquiry_type <- input$inquiry_type # creates search type)

      
      
      
      
      inFile <- input$file1
      
      if (is.null(inFile))
        return(NULL)
      
      df  <- read.csv(inFile$datapath, sep=input$sep, 
                      quote=input$quote) %>% 
        mutate(record_type = "D1",
               dob = mdy(dob),
               dob = paste0(year(dob),
                            str_pad(month(dob), 2, pad = "0"),
                            str_pad(day(dob), 2, pad = "0")),
               search_date = mdy(search_date),
               search_date = paste0(year(search_date),
                            str_pad(month(search_date), 2, pad = "0"),
                            str_pad(day(search_date), 2, pad = "0")),
               middle = str_sub(middle, start = 1, end = 1),
               record_type = record_type,
               ssn         = "",
               blank_1     = " ",
               institution_code = institution_code,
               branch_code = branch_code,
               search_date  = as.character(search_date),
               id          = as.character(id))
      
      df[is.na(df)] <- ""
      
 header <- tibble(record_type = "H1",
                       ssn = institution_code,
                       first = branch_code,
                       middle = institution_name,
                       last = date_today,
                       suffix = inquiry_type,
                       dob = "I",
                       search_date = "",
                       blank_1 = " ",
                       branch_code = "",
                       id = "",
                       institution_code = "")
      
      df <- bind_rows(header, df)
      
      
      trailer <- tibble(ssn = nrow(df) + 1,
                        record_type = "T1") %>%
        mutate(ssn = as.character(ssn),
               first = "",
               middle = "",
               last = "",
               suffix = "",
               dob = "",
               search_date = "",
               blank_1 = " ",
               branch_code = "",
               id = "",
               institution_code = "")
      
      df <- bind_rows(df, trailer)  %>%
        select(record_type, ssn, first, middle, last,  suffix, 
               dob, search_date, blank_1, institution_code, branch_code, id)
      })


```




Column {data-width=700}
-----------------------------------------------------------------------



### Instructions & Download

The purpose of this app is to remove the tedium of preparing files to submit to National Student Clearinghouse's StudentTracker service. 

To begin, create a CSV file that includes the following variables with the column names *exactly* (case sensitive) as they appear in bold below. (They don't have to be in the same order, but the columns below must be included.)

 - **first**: student's first name
 - **middle**: student's middle name or initial
 - **last**: student's last name
 - **suffix**: suffix to student's name (e.g., Jr., III)
 - **dob**: student's date of birth (mm/dd/yyyy)
 - **id**: an id code so you can link the return data back to your own data sources
 - **search_date**: search begin date (mm/dd/yyyy)
 
Next, fill in the boxes to the left according to the instructions desribed [on the Student Clearinghouses's website](http://www.studentclearinghouse.org/colleges/files/ST_ExcelInstructions.pdf). Lastly, click the `Download` button and you will have a file fully formatted to submit to the StudentTracker service.

For any questions, please contact [Arie Spirgel](mailto:spirgel@gmail.com)
 

<span style="color:grey">Code license: [MIT](https://www.r-project.org/Licenses/MIT)</span>

### Your Data Will Appear Here
```{r}

renderTable(data())

```


